---
title: "Investigating Plastic Pollution in the United States versus China Based on Cleanups from the Years 2019 and 2020"
author: "Group 4-5: Jacqueline Reed, Nasia Meimeteas, Chloe Brenner, Sofia Silvosa"
date: "April 20, 2022"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
---

```{r load-packages, message = FALSE, warning = FALSE, echo=FALSE}
library(tidyverse) 
library(tidytuesdayR)
library(tidymodels)
library(viridis)
library(knitr)
library(tidyr)
``` 

```{r plastic-pollution, message = FALSE, warning = FALSE, echo=FALSE}

plastic_pollution <- read_csv ("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv")
```

# Introduction and Data 

**Background and Significance**

Plastic pollution is one of the biggest current environmental hazards. Millions 
of animals are killed by plastic every year; marine animals are specially 
affected by plastic pollution due to ingesting or getting tangled in these 
plastic wastes.The vast majority of plastic is also produced through the use of
fossil fuels, rising gas emissions and advancing the global climate change 
crisis. It is estimated that by 2050, plastic will outweigh the amount of fish 
in the ocean.

The global plastic pollution crisis has worsened in recent years due to the 
overproduction of plastic and the growth of new plastic production facilities. 
Large corporations utilize plastic to deliver and package their products which 
has led to an over-reliance on plastic that has exacerbated this crisis. 

Companies that play a key role in the plastic pollution crisis like Coca-Cola 
and Nestle continue to implement false solutions to the crisis by utilizing 
faulty recyclable products. Companies are avoiding making real change, so
the work of the organization "Break Free from Plastic" is especially 
influential to spread awareness of the magnitude of plastic pollution. 
Analysis of this data has the potential to spread awareness of this issue
and promote corporate accountability.

**Data**

Our dataset (13,380 observations) includes information about 
plastic pollution including the country the plastic was collected in, the type 
of plastic, and the company that was the source of the plastic. The data also 
includes information about the plastic cleanup event including the number of 
events, volunteers and the total number of plastic collected.The data was 
derived from tidytuesday, & was originally collected in 2019 and 2020 from an 
audit by the NGO "Break Free from Plastic." The organization helped organize 284 
cleanups in over 50 countries; through these cleanups, Break Free From Plastic 
was able to identify the top global polluting companies and identify the most 
polluted types of plastic.

We plan to use most of the variables in the dataset to explore our research 
question. Specifically, we will focus on the following variables: country 
(country of cleanup, filtering only for the US and China), parent_company 
(source of plastic), grand_total (grand total count for all types of plastic), 
num_events (number of counting events), the 7 variables for type of plastic 
(hdpe, ldpe, o, pet, pp, ps, pvc), and year (when the plastic was collected).


**Research Questions**

When conducting this analysis, our team sought to understand:

*What are the trends of plastic pollution in the United States and China, the
two largest economies in the world?*

We hypothesize that, given the highly different governmental and social 
structures in the United States versus China, China will surpass the US in the 
amount of plastic collected in the recent years. China is considered the 
world's top manufacturing hub and the country accounts for 28.7% of total global 
manufacturing output.  

In order to examine this, we broke our analysis into three parts, each exploring 
different variables relating to plastic in these two countries. These
three variables were waste per average parent company, plastic type, and 
plastic collected per year (2019 or 2020).

This data is presented in terms of the amount of plastic collected from each 
parent company within each country. Thus, our **first research question: Is 
there a significant difference in the waste of the average parent company in 
China and the US?** 
We hypothesize that, since there is recently a large legislative push against 
pollution from businesses in the US, China will have a higher average number of 
plastic collected by a parent company. 

Secondly, there is a major problem regarding lack of proper recycling and 
organized collection of plastic and thus we wanted to look more closely at the 
data in regards to the types of plastic that was polluted. Therefore, our 
**second research question is: Within the overall reported plastic of the two 
countries, what types of plastic are most frequently found?** 
We hypothesize that the lowest amounts of plastic found will be 'HDPE' and 'PVC'
as these are the most commonly recycled plastic whereas the highest will be from 
the 'other category' as this is a "catch all" category for products that are 
not easily recycled.

Lastly, our **third research question is: How does the amount of plastic 
collected differ from 2019 to 2020?** Given that the Covid-19 pandemic hit in 
the year 2019, we predict that the amount of plastic collected in this year as 
compared to 2020 will be significantly lower due to social distancing.


# Methodology

**Exploratory Data Analysis**

We carried out the following initial analysis to discern the relationship
of grand total pollution between the countries.

```{r USChina-data, echo = FALSE, message = FALSE, warning = FALSE}
plastic_pollution1 <- plastic_pollution %>%
  filter(country == "United States of America") 
plastic_pollution2 <- plastic_pollution %>%
  filter(country == "China")
china_us_pollution <- full_join(plastic_pollution1, plastic_pollution2) %>%
  drop_na(grand_total) %>% 
  mutate(Country_binary = ifelse(country == "China", 0, 1))
```

Below is a visualization of the grand total pollution in the US and China:

```{r total-pollution, message = FALSE, warning = FALSE, echo=FALSE}
china_us_pollution %>%
  group_by(country) %>%
  filter(parent_company != "Grand Total") %>%
  ggplot(mapping = aes(x = country, y = grand_total)) + 
  geom_col() + 
  labs(title = "Grand Total Pollution in the United States and China", 
       x = "Country", y = "Grand Total Pollution")
```


Upon realizing that the data was organized by parent company, we 
created the below three visualizations to determine the highest polluting
parent companies in each country as well as the top 10 parent companies 
globally:

```{r top-10-parent-companies-in-US, message = FALSE, warning = FALSE, echo=FALSE}
china_us_pollution %>%
  filter(country=="United States of America") %>% 
  filter(parent_company != "Grand Total") %>% 
  filter(parent_company != "Unbranded") %>% 
   filter(parent_company != "null") %>% 
  arrange(desc(grand_total))  %>% 
  slice(1:10) %>% 
ggplot(mapping= aes(x= parent_company, y= grand_total)) +
  geom_col(fill='blue') + 
  coord_flip() + 
  labs(title = "Grand Total Pollution by Top 10 
       U.S Plastic Polluter Companies",
       x = "Parent Companies", y = "Grand Total Pollution")
```

```{r top-10-parent-companies-in-China, message = FALSE, warning = FALSE, echo=FALSE}
  plastic_pollution2 %>% 
  filter(parent_company != "Grand Total") %>% 
  filter(parent_company != "Unbranded") %>% 
   filter(parent_company != "null") %>% 
  arrange(desc(grand_total))  %>% 
  slice(1:10) %>% 
ggplot(mapping= aes(x= parent_company, y= grand_total)) +
  geom_col(fill='red') + 
  coord_flip() + 
  labs(title = "Grand Total Pollution by Top 10 
  China Plastic Polluter Companies",
       x = "Parent Companies", y = "Grand Total Pollution")
```

According to the Brand Audit report from 2019 organized by the
Break Free from Plastic movement, the top 10 global polluter companies were
Coca-Cola, Nestle, Pepsico, Mondelez International, Unilever, Mars, 
Proctor and Gamble, Colgate-Palmolive, Philip Morris International, and 
Perfetti. The 2020 grand total plastic amount from nine for the the U.S 
and China are found in the visual below.

```{r top-10-global-polluters, message = FALSE, warning = FALSE, echo=FALSE}
china_us_pollution %>%
  filter(parent_company != "Grand Total") %>% 
  filter(parent_company != "Unbranded") %>% 
   filter(parent_company != "null") %>% 
   filter(year == "2020") %>% 
   filter(parent_company == 'The Coca-Cola Company' | 
            parent_company == 'Procter & Gamble' | 
            parent_company == 'Pepsico'| 
            parent_company == 'Unilever' | 
            parent_company == 'Colgate-Palmolive'|  
            parent_company == 'Mondelez International' |  
            parent_company == 'Nestle' | 
            parent_company ==  'Perfetti Van Melle' | 
            parent_company ==  'Mars, Incorporated' |  
            parent_company ==  'Philip Morris International') %>% 
  ggplot(mapping=aes(x=parent_company, y= grand_total, fill=country )) +
  coord_flip() + 
  scale_fill_manual(values = c("red2","blue2")) +
   geom_col() +
   labs(title = "Grand Total Pollution of Top 10 
    Global Polluters",
       y = "Parent Companies", x = "Grand Total Pollution")
```

**RESEARCH QUESTION 1: **
*Is 
there a significant difference in the waste of the average parent company in 
China and the US?*

Based on the exploratory data analysis which revealed that the highest
polluting parent companies were different across the US and China, we first 
standardized the pollution for "the average parent company" in the US and
China and created a visualization below: 

```{r average-plastic-per-company, message = FALSE, warning = FALSE, echo=FALSE}
china_us_pollution %>% 
  group_by(country) %>% 
  filter(parent_company != "Grand Total") %>%
  filter(parent_company != "Unbranded") %>%
  filter(parent_company != "null") %>% 
  summarize(avgplasticpercompany = mean(grand_total)) %>% 
  ggplot(mapping = aes(x = country, y = avgplasticpercompany)) + 
  geom_col() + 
  labs(title = "Average Company Pollution in the United States and China", 
       x = "Country", y = "Average Company Pollution")
```

A hypothesis test was then done to check the difference in means of these two 
countries for the average amount of plastic per average parent company in each
country.

Null Hypothesis:
The average amount of plastic collected per company in China is equal to the 
mean plastic collected per parent company in the United States.
$H_0: \mu_{china} - \mu_{US} =0$

Alternative Hypothesis:
The average amount of plastic collected per parent company in China and the 
United States is not equal. 
$H_0: \mu_{china} - \mu_{US}\neq 0$

```{r research-question-1, message = FALSE, warning = FALSE, echo=FALSE}
null_dist <- china_us_pollution %>%
  filter(parent_company != "Grand Total") %>%
  filter(parent_company != "Unbranded") %>%
  filter(parent_company != "null") %>% 
  specify(response = grand_total, explanatory = country) %>%
  hypothesize(null = "independence") %>%
  generate(10000, type = "permute") %>%
  calculate(stat = "diff in means")
china_us_pollution %>% 
  group_by(country) %>% 
  filter(parent_company != "Grand Total") %>%
  filter(parent_company != "Unbranded") %>%
  filter(parent_company != "null") %>%
  summarize(diff_means = mean(grand_total)) %>% 
  kable()
visualize(null_dist) +
  shade_p_value(obs_stat = 1.01, direction = "two-sided")
get_p_value(null_dist, 1.01, "two-sided") %>% 
  kable()


```
We then used a linear regression model to explore whether country is a 
significant predictor of grand plastic total in order to further investigate 
whether there is a significant difference between the plastic pollution found 
in China compared to the United States of America.

```{r linear-regression-for-country, message = FALSE, warning = FALSE, echo=FALSE}
china_us_pollution <- china_us_pollution %>% 
filter(parent_company != "Grand Total") %>% 
mutate(dummy_country=if_else(country=='United States of America', 0,1))
m_main <- linear_reg() %>%
  set_engine("lm") %>%
  fit(grand_total ~  dummy_country , data = china_us_pollution )
 
m_main %>%
  tidy() %>% 
  kable
  glance(m_main)$r.squared %>% 
    kable

```



$$ \widehat{Grand Total} = 19.17 + 12.25~country $$ 

**RESEARCH QUESTION 2**
*Within the overall reported plastic of the two 
countries, what types of plastic are most frequently found?*

```{r rq-2-pivotlonger, message = FALSE, warning = FALSE, echo=FALSE}
longer_data <- china_us_pollution %>%
  filter(parent_company != "Grand Total") %>%  
  drop_na(hdpe, ldpe, o, pet, pp, ps, pvc) %>% 
  pivot_longer(hdpe:pvc, names_to = "type", values_to = "count_type") %>%
  select(country, type, count_type) %>% 
  group_by(country, type) %>% 
  summarize(type_number = sum(count_type)) 
```


We made a filled bar-plot visualization to investigate the distribution 
of plastic types and compare their levels between the two nations. 

```{r ggplot-research-q2, fig.align='center', message = FALSE, warning = FALSE, echo=FALSE}
ggplot(data = longer_data, aes(x= type, y= type_number, fill = country)) +
  geom_col()+
  scale_fill_manual(values = c("red2","blue2")) +
  labs(title = "Type of Plastic by Country", x = "Type of Plastic", y="Amount",
        fill= "Country")
```
Then based on the results of the visualization, we 
performed two different two-tailed t-test to determine whether
there was a significant difference in the grand total collected
for each of the leading plastic types in each country.


$H_0: \mu_{o} = \mu_{pp}$
$H_1: \mu_{o} \neq \mu_{pp}$

Null: The average type "other" plastic collected is equal to the average type 
"pp" plastic collected in the US.
Alternative: The average type "other" plastic collected is not equal to the 
average type "pp" plastic collected in the US.

```{r research-question-2-t-test, message = FALSE, warning = FALSE, echo=FALSE}
china_us_pollution%>%
  filter(country == "United States of America") %>%
  t_test(o ~ pp) %>%
  kable 
```
    
$H_0: \mu_{o} = \mu_{pet}$
$H_1: \mu_{o} \neq \mu_{pet}$

Null: The average type "other" plastic collected is equal to the average type 
"pet" plastic collected in the China.
Alternative: The average type "other" plastic collected is not equal to the 
average type "pet" plastic collected in the China.

```{r research-q2-t-test2, message = FALSE, warning = FALSE, echo=FALSE}
 china_us_pollution %>%
    filter(country == "China")%>%
    t_test(o ~ pet)%>%
    kable
```


**RESEARCH QUESTION 3**
*How does the amount of plastic 
collected differ from 2019 to 2020?*

We created a bar-plot to visualize the year over year difference in plastic
collection between 2019 and 2020.

```{r research-question-3, message = FALSE, warning = FALSE, echo=FALSE}
china_us_pollution %>% 
  group_by(country, year) %>%
  filter(parent_company != "Grand Total") %>% 
  mutate(year_categorical = ifelse(year == 2019, "2019","2020")) %>%
  mutate(grand_totalsum = sum(grand_total)) %>% 
  ggplot(mapping = aes(x = year_categorical, fill = country)) + 
  geom_bar() + 
  scale_fill_manual(values = c("red2","blue2")) +
  labs(title = "Total Company Pollution in the US and China in 2019 and 2020", 
       x = "Year", y = "Total Pollution") 
```

To test if year was a significant predictor of grand total pollution,
we performed a linear regression. 

```{r rq-3-regression, message = FALSE, warning = FALSE, echo=FALSE}
china_us_pollution <- china_us_pollution %>% 
  filter(parent_company != "Grand Total") %>% 
  mutate(dummy_year= if_else(year==2019, 0,1)) 
year_model <- linear_reg() %>%
  set_engine("lm") %>%
  fit(grand_total ~ dummy_year, data = china_us_pollution)
year_model %>%
  tidy() %>% 
  kable
  glance(year_model)$r.squared %>% 
  kable  
  
```



$$ \widehat{Grand Total} = 6.46 + 21.26~age $$ 

We then executed a multiple regression controlling for num_events and
volunteers to further explore the correlation between year and grand total.

```{r multiple-regression, message = FALSE, warning = FALSE, echo=FALSE}
options(scipen = 999)
china_us_pollution <- china_us_pollution %>% 
   filter(parent_company != "Grand Total") %>% 
  mutate(dummy_year= if_else(year==2019, 0,1)) 
year_model2 <- linear_reg() %>%
  set_engine("lm") %>%
  fit(grand_total ~ dummy_year + num_events + volunteers, 
      data = china_us_pollution)
year_model2 %>%
  tidy() %>% 
  kable
  glance(year_model2)$r.squared %>% 
  kable
```

$$ \widehat{Grand Total} = 3.96 +15.25~year +0.15~events -0.004~volunteers $$ 



# Results

**Research question 1:** 

The first visualization revealed that the average parent company in China
released more total pollution than the average US company. Based on this
information, we performed a hypothesis test to further explore the relationship.

The observed statistic was a difference in means of 1.01. 
The p-value that was obtained is 0.4122 This value is well over 0.05. This means 
that we fail to reject the null hypothesis. In the context of this situation, we 
cannot reject the that the difference between the average 
amount of plastic collected per company for China and the US is the same. 

The intercept in our linear
regression model is 19.17, meaning that if there was 
no 'country', we would expect the grand plastic total to be 19.17 units. 
In the context of this model, this intercept value is not applicable.
If the grand total plastic count was collected in China, we would expect it 
the plastic count to be 12.25 units less than in the United States. 
0.12% percent of the variance in grand plastic total amount is explained by 
country.

However, our explanatory variable of 'country' is not significant at the 
$a$ = 0.05 level, meaning that this predictor variable is not statistically
significant. Thus, we do not have enough evidence to support the claim 
that the country in which the plastic pollution was collected predicts
total grand total plastic amount.

**Research question 2:**
Based on the visualization created, it is clear that the most prevalent type of 
plastic categorized in collection was the type “o”, which stood for other in the
dataset. This is any type of plastic gathered that did not fall under the other 
categories given. Almost all values are zero besides the “o” category and a 
small number of plastics collected in the “pet” category, which is polyester.

Polyester has the highest value for China (not counting the
other category), it does not have the highest value for the US. In the US, the 
most prevalent type of plastic was “pp”, or Polypropylene. Closely under the 
amount of polypropylene collected in the US was the amount of polyester. The 
rest of the plastic types were found with much lower frequency than these two 
types were.  Low density polyethylene and Polystyrene were collected in similar 
amounts. Right beneath these two types of plastic was high density polyethylene. 
Last, the least prevalent type was PVC.

Despite the differing bar distributions, neither the difference between the 
number of "pp" and "other" collected in the US and "other" and "pet" collected
China are significant different at $\alpha=0.05$ level (p=0.165, p= 0.565). 



**Research question 3:**

Based on our visual, we can conclude that that in 2019 there was a greater 
amount of plastic grand total count for both China and the United States. 


According to our linear regression model, if the year was 2020, 
we would expect the grand total amount 
to be 21.26 units higher than if it was 2019.  0.32% of the variance
in grand plastic total count is explained by the
year the plastic total was collected.

The year predictor was significant at the $\alpha$ = 0.05 level (p=0.017),so 
based on this linear regression model the relationship between
year and grand total plastic count is statistically significant. However, in
order to ensure that we are avoiding a Type 1 error, we decided to also
run a multiple regression model controlling for number of volunteering
events and volunteer count to investigate whether year still plays a significant
predictor role.

According to our multiple regression model, for every increase in
volunteering events the plastic grand total count 
increases by 0.15, all else held constant. For every increase in volunteers
the plastic grand total count decreases by 0.004, all else held constant. 
Finally, for our year variable we can interpret our main effect as if the year
was 2020, we would expect the grand total amount to be 15.25 units higher than 
if it was 2019.

As none of the predictors were below the $\alpha$ = 0.05 level, we do 
not have evidence to demonstrate these these variables predict our
explanatory variable of total plastic grand total. 0.49% of the variance 
in grand plastic total count is explained by year, volunteers,
and volunteering events.


# Discussion 

**Conclusions:**

RQ1: According to our methodology, the difference in 
grand total plastic waste from the average parent company in China and 
the US is not significantly different. The evidence for this conclusion 
is the hypothesis test investigating the difference in mean grand total waste
is not significant at the $\alpha=0.05$ level. This conclusion contradicts
our original hypothesis that Chinese parent companies would emit
greater levels of pollution due to US legislation focused on limiting 
plastic waste. However, the validity of this 
data may be called into question due to the unequal number of 
collection events.

RQ2: It is reasonable that "o" is the highest ranking column 
in the visualization for both the US and China 
(as visible when looking at the proportions reflected in 
the chart) since it likely encompasses a lot of different plastic types that 
were not identified by volunteers. When considering the plastic collected that 
actually fit one of the coded types, it is clear that there is very little data 
for China present. By contrast polyester was the most frequently
collected category for China, which is not surprising given the context 
that there is an established polyester surplus in China (https://www.fibre2fashion.com/industry-article/7406/surplus-polyester-
production-in-china). However, we conclude that the differences
between the number of items collected between the top two plastic types
for each country are not significant. Therefore, rather our results
do not suggest that targeting a specific type of plastic in regulations
will be more effective in reducing overall pollution. 


RQ3: Based on the visualization from research question 3, more plastic was
collected in both nations in 2019 compared to 2020. 
Perhaps, this is a result of a decrease in plastic pollution in 2020. However, 
this is likely not the case as shown with recent trends of plastic pollution as 
outlined by the Breaking Free from Plastic report. This difference in 2019 to 
2020 plastic count is more likely a result of the decrease of volunteering clean 
up events because of the COVID-19 pandemic.

In addition, the linear regression model to investigate whether 
year predicts grand total plastic indicated that it is significantly 
correlated as evidenced by the 0.017 p value. However, due to an observed
difference in the number of collection events  and volunteers
between the two countries during initial exploratory data analysis, we ran a 
multiple regression to control for these differences. The regression 
demonstrated that when number of events and volunteers were controlled, 
year was not significant at the alpha level (p= 0.204). Furthermore,
neither the number of events or volunteers were significant
all else held constant. So, in the context of the data, we fail to reject the 
null hypothesis that any of these three predictors significantly affects
grand total of plastic found. The difference between
from the linear and multiple regressions shows that grand total is not
correlated with year despite that collection in 15.25 units higher year
over year. 

This regression result is surprising because 2020 was a pandemic year. 
This has two implications: firstly that single use plastics increased 
in the pandemic as online consumption increased. Secondly, that 
less data was likely able to be collected in 2020 maybe because access
to certain areas and social gatherings was limited. As such, we 
hypothesized that year would be significant predictor for grand total, but
it was not. Despite that number of events and volunteers were controlled, 
it is reasonable to assume that the reality of the increase in plastic s
should have increased the grand total significantly. 

Overall: We used these three research questions to investigate our
overarching question about the plastic pollution trends in
US and China. We found that the average parent company in each
country did not produce significantly different amounts of pollution.
Furthermore,the differences in grand total plastic
between years and plastic types were negligible. As such, our analysis
indicates that the differences between the variables studied in each
question are largely insignificant.However, there is 
still much to be investigated about this topic as plastic pollution
increasingly harms human, ocean, and ecosystem health. 

**Limitations:**
There were a few significant limitations to our data analysis. 
The first was the difference in the number of collection events both 
between China and the US and 2019 and 2020. These differences are reasonable
as 2020 was a pandemic year, and with China's stringent regulations, 
holding frequent plastic collections where a large group would have to 
gather was not feasible. Furthermore, the organization Break-Free from Plastic,
where the data comes from, has 111 core collection partners in the United
States whereas they only have 1 core collection partner in China. Therefore,
there are bound to be more US collection data points. Given these
realities, the findings related to higher plastic pollution in the US 
may be artificially inflated. 

Another key limitation was that the data itself was grouped by parent company, 
and for the majority, the parent companies in the US differ from those 
in China. As such, we focused on the differences between the average
parent company in China vs the US instead of another more general variable
such as difference in total plastic per province/state or collected
per month. 

**Further Research:**

In future analysis, we could base our analysis on 
background research about plastic in the China and the US. For example, 
what plastics have been most commonly found in each country
or which plastic types are most harmful. Using statistics to further research
existing conclusions is important.

As the original plastic_pollution data set is very extensive with 13,380
observations, in the future it would be fascinating to analyze the
differences to the same research questions we asked (about parent
companies, plastic types, and year-over-year differences) but for all or more
of the countries in the data set. We had difficulty working with this much
data, and we encountered impassable difficulties in grouping by country
so decided to do a simplified comparison of China and the US, the world's 
greatest polluters. But, future research should focus on countries or
regions more broadly. 


**References**

- https://www.breakfreefromplastic.org/2019/10/23/brand-audit-report-2019-press-
release/

- https://globalupside.com/top-10-manufacturing-countries-in-the-world/#:~:text=
China%20makes%20up%2028.7%25%20of,China%20claimed%20the%20top%20position 

- https://www.nationalgeographic.com/environment/article/plastic-pollution

- https://www.breakfreefromplastic.org/globalbrandauditreport2019/

- https://www.biologicaldiversity.org/campaigns/ocean_plastics/#:~:text=Plastics
%20pollution%20has%20a%20direct,or%20getting%20entangled%20in%20it.

- https://scc.ms.unimelb.edu.au/resources-list/simple-r-scripts-for-analysis/r-scr
ipts
